CrypTool Script \\ Mathematics and Cryptography

### 2003-03-07 script/en/chapters/aboutcryptool.tex
no message

\parskip 4pt
\vskip + 30 pt
{
In this CrypTool script you will find predominantly mathematically oriented
information for using encryption procedures. The main chapters have been written
by {\em various authors} and are therefore independent from one another. At the end of
most chapters you will find literature and web links.

@@

Whereas the program CrypTool teaches you how to use cryptography in practice, the
script provides those interested in the subject with a deeper understanding of
the mathematical algorithms used -- trying to do it didactically as good as possible.

The {\em authors} Bernhard Esslinger, Bartol Filipovic, Henrik Koy,
Roger Oyono and J\"org Cornelius Schneider
would like to take this opportunity to thank their colleagues in the company and
at the universities of Frankfurt, Giessen, Siegen and Karlsruhe. They particularly thank
Dr. Peer Wichmann from the Karlsruhe computer science research centre
(Forschungszentrum Informatik, FZI) for his down-to-earth support.

@@

\vskip +7 pt \noindent
You will find the current version of CrypTool under \newline
  \href{http://www.CrypTool.de}{\texttt{http://www.cryptool.de}},~
  \href{http://www.cryptool.com}{\texttt{http://www.cryptool.com}}~ or~
  \href{http://www.cryptool.org}{\texttt{http://www.cryptool.org}}.
{
\vskip + 7 pt \noindent
The contact persons for this free tool are listed in the readme file for
CrypTool.}
}

>>>>

% ............................................................................
%                 TEXT DER 1. SEITE
% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

\parskip 4pt
\vskip + 30 pt
{
In this CrypTool\index{CrypTool} script you will find predominantly mathematically oriented
information on using cryptographic procedures. The main chapters have been written
by {\em various authors} and are therefore independent from one another. At the end of
most chapters you will find literature and web links.

@@

Whereas the program CrypTool\index{CrypTool} teaches you how to use cryptography in practice, the
script provides those interested in the subject with a deeper understanding of
the mathematical algorithms used -- trying to do it didactically as good as possible.

The {\em authors} Bernhard Esslinger, Mathias B"uger,Bartol Filipovic, 
Henrik Koy, Roger Oyono and J\"org Cornelius Schneider
would like to take this opportunity to thank their colleagues in the company and
at the universities of Frankfurt, Gie\ss en, Siegen and Karlsruhe. They particularly thank
Dr. Peer Wichmann from the Karlsruhe computer science research centre
(Forschungszentrum Informatik, FZI) for his down-to-earth support.

@@

\vskip +7 pt \noindent
You will find the current version of CrypTool\index{CrypTool} under \newline
  \href{http://www.CrypTool.de}{\texttt{http://www.cryptool.de}},~
  \href{http://www.cryptool.com}{\texttt{http://www.cryptool.com}}~ or~
  \href{http://www.cryptool.org}{\texttt{http://www.cryptool.org}}.
{
\vskip + 7 pt \noindent
The contact persons for this free tool are listed in the readme file for
CrypTool.}
}

### 2003-03-07 script/en/chapters/cryptomethods.tex
be comments
no message

\newpage
\section{Encryption procedures}

\subsection{Encryption}

@@

\subsubsection{Symmetric encryption}

@@

\subsubsection{Asymmetric encryption}

@@

If Alice wants to communicate with Bob, then she finds Bob's public key in the
directory and uses it to encrypt her message to him. She then sends this
ciphertext to Bob, who is then able to decrypt it again using his secret key. As
only Bob knows his secret key, only he can decrypt messages addressed to him.
Even Alice who sends the message cannot restore plaintext from the (encrypted)
message she has sent. Of course, you must first ensure that the public key
cannot be used to derive the private key.\par \vskip + 3pt

@@

\newpage
\subsubsection{Hybrid procedures}
\index{Hybrid procedures}
In order to benefit from the advantages of symmetric and asymmetric techniques
together, hybrid procedures are usually used (for encryption) in practice.\par \vskip + 3pt

@@

\subsubsection{Further details}

Beside information you can find in many books and on a lot of websites the online help
of CrypTool also offers very many details about the symmetric and asymmetric encryption methods.

\begin{thebibliography}{99999}
\addcontentsline{toc}{subsection}{Literatur}
\bibitem[Schmeh2001]{Schmeh2001}  \index{Schmeh 2001}
        Klaus Schmeh, \\
        Kryptografie und Public-Key Infrastrukturen im Internet, dpunkt.verlag, 2. akt. und erw. Auflage 2001. \\
        A considerable, up-to-date, good reading book, which considers practical problems, like standardisation or
        real existing software. Currently published only in German language.
\end{thebibliography}

>>>>

% ..............................................................................
%                V E R S C H L U E S S E L U N G S V E R F A H R E N
% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

\newpage
\section{Encryption procedures}
\hypertarget{Kapitel_1}{}
(Bernhard Esslinger, besslinger@web.de, May 1999, Updates Dec. 2001, Feb. 2003)

% --------------------------------------------------------------------------
\subsection{Encryption}

@@

% --------------------------------------------------------------------------
\subsubsection[Symmetric encryption]
{Symmetric encryption\footnotemark}
\footnotetext{%
With CrypTool\index{CrypTool} v1.3 you can execute the following modern
symmetric encryption algorithms 
(using the menu ``Crypt \textbackslash Symmetric''): \\
IDEA, RC2, RC4, DES (ECB), DES (CBC), Triple-DES (ECB), Triple-DES (CBC),
MARS (AES candidate), RC6 (AES candidate), Serpent (AES candidate), 
Twofish (AES candidate), Rijndael (official AES algorithm).
}

@@

More details about the AES algorithms can be found within the Online help of CrypTool\index{CrypTool} (in the index see head-word {\em AES} and then the help pages {\em AES candidates}, {\em The AES Winner Rijndael} and {\em The Rijndael encryption algorithm}).

% --------------------------------------------------------------------------
\subsubsection{New results about cryptanalysis of AES}

Below you will find some results, which have recently called into question the security of the AES algorithm -- from our point of view these doubts practically still remain unfounded. The following information is based on the original papers and the articles \cite{Wobst-iX2002} and \cite{Lucks-DuD2002}.

AES with a minimum key length of 128 bit is still in the long run sufficiently secure against brute-force attacks - as long as the quantum computers weren't powerful enough. When announced as new standard AES was immune against all known crypto attacks, mostly based on statistical considerations and earlier applied to DES: using pairs of clear and cipher texts expressions are constructed, which are not completely at random, so they allow conclusions to the used keys. These attacks required unrealistically large amounts of intercepted data.

Cryptanalysts already label methods as ``academic success'' or as ``cryptanalytic attack'' if they are theoretically faster than the complete testing of all keys (brute force analysis). In the case of AES with the maximal key length (256 bit) exhaustive key search on average needs $2^{255}$ encryption operations. A cryptanalytic attack needs to be better than this. At present between $2^{75}$ and $2^{90}$ encryption operations are estimated to be performable for organizations, for example a security agency.

In their 2001-paper Ferguson, Schroeppel and Whiting \cite{Ferguson2001} presented a new method of symmetric codes cryptanalysis: They described AES with a closed formula (in the form of a continued fraction) which was possible because of the "relatively" clear structure of AES. This formula consists of around 1000 trillion terms of a sum - so it does not help concrete practical cryptanalysis. Nevertheless -curiosity in the academic community awakened. It was already known, that the 128-bit AES could be described as an over-determined system of about 8000 quadratic equations (over an algebraic number field) with about 1600 variables (some of them are the bits of the wanted key) -- equation systems of that size are in practice not solvable. This special equation system is relatively sparse, so only very few of the quadratic terms (there are about 1,280,000 are possible quadratic terms in total) appear in the equation system.

The mathematicians Courois and Pieprzyk \cite{Courtois2002} published a paper in 2002, which got a great deal of attention amongst the crypto community: The pair had further developed the XL-method (eXtended Linearization), introduced at Eurocrypt 2000 by Shamir et al., to create the so called XSL-method (eXtended Sparse Linearization). The XL-method is a heuristic technique, which in some cases manages to solve big non-linear equation systems and which was till then used to analyze an asymmetric algorithm (HFE).  The innovation of Courois and Pieprzyk was, to apply the XL-method on symmetric codes: the XSL-method can be applied to very specific equation systems. A 256-bit AES could be attacked in roughly $2^{230}$ steps. This is still a purely academic attack, but also a direction pointer for a complete class of block ciphers. The major problem with this attack is that until now nobody has worked out, under what conditions it is successful: the authors specify in their paper necessary conditions, but it is not known, which conditions are sufficient.
There are two very new aspects of this attack: firstly this attack is not based on statistics but on algebra. So attacks seem to be possible, where only very small amounts of cipher text are available. Secondly the security of a product-algorithm does not exponentially increase with the number of rounds.

Currently there is a large amount of research in this area: for example Murphy and Robshaw presented a paper at Crypto 2002 \cite{Robshaw2002a}, which could dramatically improve cryptanalysis: the burden for a 128-bit key was estimated at about $2^{100}$ steps by describing AES as a special case of an algorithm called BES (Big Encryption System), which has an especially "round" structure. But even $2^{100}$ steps are beyond what is achievable in the foreseeable future. Using a 256 bit key the authors estimate that a XSL-attack will require $2^{200}$ operations.

More details can be found at: \\
\href{http://www.cryptosystem.net/aes}{\texttt{http://www.cryptosystem.net/aes}} and \\ 
\href{http://www.minrank.org/aes/}{\texttt{http://www.minrank.org/aes/}}

So for 256-AES the attack is much more effective than brute-force but still far away from any computing power which could be accessible in the short-to-long term. The discussion is very controversial: Don Coppersmith (one of the inventors of DES) for example queries the practicability of the attack because XLS would provide no solution for AES \cite{Coppersmith2002}. This concludes that then the optimization of Murphy and Robshaw \cite{Robshaw2002b} would not work.

% --------------------------------------------------------------------------
\subsubsection{Asymmetric encryption}

@@

If Alice wants to communicate with Bob, then she finds Bob's public key 
in the directory and uses it to encrypt her message to him. She then sends
this cipher text to Bob, who is then able to decrypt it again using his 
secret key. As only Bob knows his secret key, only he can decrypt 
messages addressed to him.
Even Alice who sends the message cannot restore plaintext from the (encrypted)
message she has sent. Of course, you must first ensure that the public key
cannot be used to derive the private key.\par \vskip + 3pt

@@

% --------------------------------------------------------------------------
\newpage
\subsubsection[Hybrid procedures]
{Hybrid procedures\footnotemark}
\footnotetext{%
Within CrypTool\index{CrypTool} v1.3 you can get a visualization of this
technique using the menu ``Crypt \textbackslash Hybrid Demonstration'': 
this dialogue shows the single steps and its dependencies with concrete
numbers.
}\index{Hybrid procedures}

In order to benefit from the advantages of symmetric and asymmetric techniques
together, hybrid procedures are usually used (for encryption) in practice.
\par \vskip + 3pt

@@

% --------------------------------------------------------------------------
\subsubsection{Further details}

Beside information you can find in many books and on a lot of websites the 
online help of CrypTool\index{CrypTool} also offers very many details about
the symmetric and asymmetric encryption methods.

% --------------------------------------------------------------------------
\begin{thebibliography}{99999}
\addcontentsline{toc}{subsection}{Bibliography}

\bibitem[Schmeh2003]{Schmeh2003}  \index{Schmeh 2003}
        Klaus Schmeh, \\
        {\em Cryptography and Public Key Infrastructures on the Internet}, 
	John Wiley \& Sons Ltd., Chichester 2003. \\
        A considerable, up-to-date, good reading book, which also 
	considers practical problems, like standardisation or
        real existing software.

\bibitem[Coppersmith2002]{Coppersmith2002}  \index{Coppersmith 2002}
        Don Coppersmith, \\
        {\em Re: Impact of Courtois and Pieprzyk results}, 
	2002-09-19, AES Discussion Groups at \\
        \href{http://aes.nist.gov/aes/}
        {\texttt{http://aes.nist.gov/aes/}}

\bibitem[Courtois2002]{Courtois2002}  \index{Courtois 2002}
        Nicolas Courtois, Josef Pieprzyk, \\
        {\em Cryptanalysis of Block Ciphers with Overdefined Systems of Equations}, 
	received 10 Apr 2002, last revised 9 Nov 2002,
	A different version, so called compact version of the first XSL attack,
	was published in Asiacrypt Dec 2002, \\
        \href{http://eprint.iacr.org/2002/044}
        {\texttt{http://eprint.iacr.org/2002/044}}

\bibitem[Ferguson2001]{Ferguson2001}  \index{Ferguson 2001}
        Niels Ferguson, Richard Schroeppel, Doug Whiting, \\
        {\em A simple algebraic representation of Rijndael}, 
	Draft 2001/05/1, \\
        \href{http://www.xs4all.nl/~vorpal/pubs/rdalgeq.html}
        {\texttt{http://www.xs4all.nl/\~{}vorpal/pubs/rdalgeq.html}}

\bibitem[Lucks-DuD2002]{Lucks-DuD2002}  \index{Lucks 2002}
        Stefan Lucks, R"udiger Weis, \\
        {\em Neue Ergebnisse zur Sicherheit des Verschlüsselungsstandards AES}, 
	in DuD Dec. 2002.

\bibitem[Robshaw2002a]{Robshaw2002a}  \index{Robshaw 2002}
        S.P. Murphy, M.J.B. Robshaw, \\
        {\em Essential Algebraic Structure within the AES}, 
	June 5, 2002, Crypto 2002,  \\
        \href{http://www.isg.rhul.ac.uk/\~{}mrobshaw/rijndael/rijndael.html}
        {\texttt{http://www.isg.rhul.ac.uk/~mrobshaw/rijndael/rijndael.html}}

\bibitem[Robshaw2002b]{Robshaw2002b}  \index{Robshaw 2002}
        S.P. Murphy, M.J.B. Robshaw, \\
        {\em Comments on the Security of the AES and the XSL Technique}, 
	September 26, 2002, \\
        \href{http://www.isg.rhul.ac.uk/\~{}mrobshaw/rijndael/rijndael.html}
        {\texttt{http://www.isg.rhul.ac.uk/~mrobshaw/rijndael/rijndael.html}}

\bibitem[Wobst-iX2002]{Wobst-iX2002}  \index{Wobst 2002}
        Reinhard Wobst, \\
        {\em Angekratzt - Kryptoanalyse von AES schreitet voran}, 
	in iX Dec. 2002, \\
	plus the reader's remark by Johannes Merkle in iX Feb. 2003.

\end{thebibliography}

### 2003-03-07 script/en/chapters/digitalsignatures.tex
be fixes
no message

@@

The aim of digital signatures is to guarantee the following two points:
\begin{itemize}
 \item User authenticity: \\
      \index{Authenticity!user} It can be checked whether a message really does
come from a particular person.
 \item Message integrity: \\
      \index{Message integrity} It can be checked whether the message has been
changed (on route).
\end{itemize}

An asymmetric technique is used again (see encryption procedures). Participants
who wish to generate a digital signature for a document must possess a pair of
keys. They use their secret key to generate signatures and the recipient uses
the sender's public key to verify whether the signature is correct. As before,
it must be impossible to use the public key to derive the secret key.

@@

The document is sent to the recipient together with the signature. The recipient
can then use the sender's public key, the document and the signature to
establish whether or not the signature is correct. In practice, however, the
procedure we have just described has a decisive disadvantage. The signature is
approximately as long as the document itself. To prevent an unnecessary increase
in data traffic, and also for reasons of performance\index{Performance}, we use a cryptographic
hash function.

@@

        \bibitem[FIPS186]{FIPS186} U.S. Departement of
          Commerce/N.I.S.T.\\\index{FIPS186}
          Entity authentication using public key cryptography. February 18, 1997.
\end{thebibliography}

>>>>

% ...........................................................................
%                  D I G I T A L E  S I G N A T U R E N
% ...........................................................................

@@

\vspace{12pt}
The aim of digital signatures is to guarantee the following two points:
\begin{itemize}
 \item User authenticity: \index{Authenticity!user} \\
      It can be checked whether a message really does
come from a particular person.
 \item Message integrity:  \index{Message integrity} \\
      It can be checked whether the message has been
changed (on route).
\end{itemize}

An asymmetric technique is used again (see encryption procedures). Participants
who wish to generate a digital signature for a document must possess a pair of
keys. They use their secret key to generate signatures and the recipient uses
the sender's public key to verify whether the signature is correct. As before,
it must be impossible to use the public key to derive the secret key\footnote{%
With CrypTool\index{CrypTool} v1.3 you can also generate and check 
digital signatures: \\
using the submenus of the main menu ``Digital Signatures'' or \\
using menu {\bf Indiv. Procedures \textbackslash RSA Demonstration 
\textbackslash Signature Demonstration}.
}.

@@

The document is sent to the recipient together with the signature. The recipient
can then use the sender's public key, the document and the signature to
establish whether or not the signature is correct. In practice, however, the
procedure we have just described has a decisive disadvantage. The signature is
approximately as long as the document itself. To prevent an unnecessary increase
in data traffic, and also for reasons of performance\index{Performance}, we use 
a cryptographic hash function\footnote{%
%xxxxxxxxxx Menüeinträgsbezeichnungen ok?
Hash functions are implemented within CrypTool\index{CrypTool} v1.3.04 at
several places: \\
using menu  ``Indiv. Procedures \textbackslash Hash'' you have the possibility
\begin{itemize}\addtolength{\itemsep}{-1.0\baselineskip} % um den Abstand zu verringern bebe
\item to apply one of 6 hash functions to the content of the current window, \\
\item to calculate the hash value of a file, \\
\item to test, how changes to a text change the according hah value and \\
\item to perform a simulation, how digital signatures could be attacked 
      by a targeted search for hash value collisions.
\end{itemize}
}
.

@@

        \bibitem[FIPS186]{FIPS186} U.S. Department of
          Commerce/N.I.S.T.\\\index{FIPS186}
          Entity authentication using public key cryptography. February 18, 1997.
\end{thebibliography}

### 2003-03-07 script/en/chapters/ellipticcurves.tex
be fixes
no message

@@

\subsection{Elliptic curce cryptography --- a high-performance substitute for RSA?}\index{Performance}\label{ECAlternative}

@@

For secure curves, the parameter $p$ determines how long it takes
to solve the ECDLP on this curve. The larger the parameter $p$,
the longer it takes to solve the problem. Experts recommend a bit
length of over $200$ bits for the parameter $p$. This makes it
clear why elliptic curves are so interesting for cryptography.
Because the parameter $p$ also determines the time required to
perform the signature/encryption procedure when using elliptic
curves in cryptography. The time taken to generate a pair of keys
also depends on $p$. Thus, small values (few bits) are desirable
here (in order to minimise the run times for the procedures);
however, the required security must still be maintained. For
example, with a length of $200$ bits for $p$, a {\em good}
elliptic curve is just as secure as an \index{RSA!module} RSA
module of over $1024$ bits in length (at least according to the
current state of research). The reason for this is that the
quickest algorithms for solving the {\em elliptic curve discrete
logarithm} problem have an exponential run time --- unlike the
subexponential run times that the best factorisation algorithms
currently have (number sieve, quadratic sieve or factorisation
with elliptic curves). Therefore, the parameters for cryptographic
procedures based on the problem of {\em factorising whole numbers}
must be greater than the parameters for cryptographic procedures
based on the ECDLP problem.

@@

There are factorisation algorithms based on elliptic curves\footnote{In 1987 H.W. Lenstra published
a factorisation algorithm, based on elliptic curves (see \cite{Lenstra1987}).
The biggest compound number curently factorised with elliptic curves is the number $ 628^{59}-1, $ which has 55 decimal digits. It was
found Oct. 6th, 2001 by M. Izumi (See \hyperlink{Lenstra2}{ECMNET}\index{ECMNET}).
}. More precisely,
these procedures exploit the fact that elliptic curves can be defined over
${\mathbb Z}_n$ ($n$ composite number). Elliptic curves over ${\mathbb Z}_n$ do
not form a group, because not every point on such an elliptic curve has an
inverse point. This is connected with the fact that - if $n$ is a composite
number - there exist elements in ${\mathbb Z}_n$ that do not have an inverse
with respect to multiplication mod $n$. In order to add two points on an
elliptic curve over ${\mathbb Z}_n$, we can calculate in the same way as on
elliptic curves over ${\mathbb Z}_p$. Addition of two points (on an elliptic
curve over ${\mathbb Z}_n$), however, fails if and only if a factor of $n$ has
been found. The reason for this is that the procedure for adding points on
elliptic curves gives elements in ${\mathbb Z}_n$ and calculates the inverse
elements for these (with respect to multiplication mod $n$) in ${\mathbb Z}_n$.
The extended \index{Euclidean algorithm} Euclidean algorithm is used here. If
the addition of two points (that lie of an elliptic curve over ${\mathbb Z}_n$)
gives an element in ${\mathbb Z}_n$ that does not have an inverse element in
${\mathbb Z}_n$, then the extended Euclidean algorithm delivers a genuine factor
of $n$.

@@

CrypTool uses elliptic curves for the digital signature function.

@@

\begin{enumerate}
   \item Certicom – Online Tutorial, \index{Certicom}\\
                \href{http://www.certicom.com/resources/ecc_tutorial/ecc_tutorial.html}{\texttt{http://www.certicom.com/resources/ecc\_tutorial/ecc\_tutorial.html}}
        \item IEEE P1363

>>>>

% ...........................................................................
%                  E L L I P T I S C H E  K U R V E N
% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

@@

\subsection{Elliptic curve cryptography --- a high-performance substitute for RSA?}\index{Performance}\label{ECAlternative}

@@

For secure curves, the parameter $p$ determines how long it takes
to solve the ECDLP on this curve. The larger the parameter $p$,
the longer it takes to solve the problem. Experts recommend a bit
length of over $200$ bits for the parameter $p$. This makes it
clear why elliptic curves are so interesting for cryptography.
Because the parameter $p$ also determines the time required to
perform the signature/encryption procedure when using elliptic
curves in cryptography. The time taken to generate a pair of keys
also depends on $p$. Thus, small values (few bits) are desirable
here (in order to minimise the run times for the procedures);
however, the required security must still be maintained. For
example, with a length of $200$ bits for $p$, a {\em good}
elliptic curve is just as secure as an \index{RSA!module} RSA
module of over $1024$ bits in length (at least according to the
current state of research). The reason for this is that the
quickest algorithms for solving the {\em elliptic curve discrete
logarithm} problem have an exponential run time --- unlike the
sub-exponential run times that the best factorisation algorithms
currently have (number sieve, quadratic sieve or factorisation
with elliptic curves). Therefore, the parameters for cryptographic
procedures based on the problem of {\em factorising whole numbers}
must be greater than the parameters for cryptographic procedures
based on the ECDLP problem.

@@

There are factorisation algorithms based on elliptic curves
\footnote{In 1987 H.W. Lenstra published
a factorisation algorithm, based on elliptic curves (see \cite{Lenstra1987}).
The biggest compound number currently factorised with elliptic curves is 
the number $ 628^{59}-1, $ which has 55 decimal digits. It was
found Oct. 6th, 2001 by M. Izumi 
(See \hyperlink{Lenstra2}{ECMNET}\index{ECMNET}).
}. 
More precisely, these procedures exploit the fact that elliptic curves can
be defined over ${\mathbb Z}_n$ ($n$ composite number). Elliptic curves 
over ${\mathbb Z}_n$ do not form a group, because not every point on such 
an elliptic curve has an inverse point. This is connected with the fact 
that - if $n$ is a composite
number - there exist elements in ${\mathbb Z}_n$ that do not have an inverse
with respect to multiplication mod $n$. In order to add two points on an
elliptic curve over ${\mathbb Z}_n$, we can calculate in the same way as on
elliptic curves over ${\mathbb Z}_p$. Addition of two points (on an elliptic
curve over ${\mathbb Z}_n$), however, fails if and only if a factor of $n$ has
been found. The reason for this is that the procedure for adding points on
elliptic curves gives elements in ${\mathbb Z}_n$ and calculates the inverse
elements for these (with respect to multiplication mod $n$) in ${\mathbb Z}_n$.
The extended \index{Euclidean algorithm} Euclidean algorithm is used here. If
the addition of two points (that lie of an elliptic curve over ${\mathbb Z}_n$)
gives an element in ${\mathbb Z}_n$ that does not have an inverse element in
${\mathbb Z}_n$, then the extended Euclidean algorithm delivers a genuine factor
of $n$.

@@

CrypTool\index{CrypTool} uses elliptic curves for the digital signature function.

@@

\begin{enumerate}
   \item Certicom Online Tutorial, \index{Certicom}\\
                \href{http://www.certicom.com/resources/ecc_tutorial/ecc_tutorial.html}{\texttt{http://www.certicom.com/resources/ecc\_tutorial/ecc\_tutorial.html}}
        \item IEEE P1363

### 2003-03-07 script/en/chapters/introduction.tex
no message

@@

This script is delivered together with CrypTool.

@@

Since then CrypTool has been used for education in companies and
universities and several universities help to develop further project.

Because the articles in this script are largely self-contained, this script can also
be read independantly of CrypTool.

@@

Additionally we aimed to describe the newest results, as well as their
origins, about primes and factorization in order to be highly topical.

>>>>

% ............................................................................
%                         E I N F Ü H R U N G
% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

@@

This script is delivered together with CrypTool\index{CrypTool}.

@@

Since then CrypTool\index{CrypTool} has been used for education in companies and
universities and several universities help to develop further project.

Because the articles in this script are largely self-contained, this script can also
be read independently of CrypTool\index{CrypTool}.

@@

Additionally we aimed to describe the newest results, as well as their
origins, concerning primes, factorization and cryptanalysis of AES, in order to be highly topical.
\\

At this point I'd like to thank explicitely 3 persons who espcially 
contributed to CrypTool. Without their talents and engagement 
CrypTool would not be what it is today:
\begin{itemize}
   \item Mr. Henrik Koy
   \item Mr. Joerg-Cornelius Schneider and
   \item Dr. Peer Wichmann.
\end{itemize}
Also I want to thank all the ones not named here for their engagement (mostly
performed in their spare time).
\\

Bernhard Esslinger

Frankfurt, March 2003
